---
title: |
  |  \Large SIR cooties
author: | 
  | \large Gregory J. Matthews \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
abstract: |
  | Cooties. \vspace{2mm}
  | *Keywords*: Cooties
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
theme_set(theme_minimal())
library(Lahman)
```

\newpage

# Introduction {#sec:intro}
SIR models are commonly used as a simple model for epidemics.  
SIR Lit review

## Zombies


# Methods {#sec:methods}
These models can be expressed as differential equations with the solution describing the trajectories of the susceptible-infected-recovered groups.  The traditional SIR model is described by the following differential equations: 

$$
\frac{dS}{dt} = \frac{-\beta SI}{N}\\
\frac{dI}{dt} = \frac{\beta SI}{N} - \gamma I\\
\frac{dR}{dt} = \gamma I
$$

with $S$, $I$, and $R$ representing the number of sick, infected, and recovered individuals, $N$ is the total number of people in the system such that $S + I + R = N$ and $\gamma$ and $\beta$ are the rates (?) of recovery and infection, respectively.  

# Results {#sec:results}

# Conclusion {#sec:conclusion}

# Acknowledgements {-}

# Supplementary Material {-}

All code for reproducing the analyses in this paper is publicly available at https://github.com/gjm112/HockeyEnforcers

# References
