---
title: |
  |  \Large SIR cooties
author: | 
  | \large Gregory J. Matthews \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
abstract: |
  | Cooties. \vspace{2mm}
  | *Keywords*: Cooties
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
theme_set(theme_minimal())
library(Lahman)
```

\newpage

# Introduction {#sec:intro}
SIR models (@weiss2013sir) are commonly used as a simple model for epidemics.  

Covid: @ellison2020implications, @cooper2020sir, @kudryashov2021analytical

Measles: @bjornstad2002dynamics, @arsal2020short, @allen1991discrete


SIR Lit review

## Zombies
Many 

@reehl2013zombie, @simeonov2022humans, @witkowski2013bayesian, @libal2023transition, @alemi2015you

# Methods {#sec:methods}
These models can be expressed as differential equations with the solution describing the trajectories of the susceptible-infected-recovered groups.  The traditional SIR model is described by the following differential equations: 

$$
\frac{dS}{dt} = \frac{-\beta SI}{N}\\
$$
$$
\frac{dI}{dt} = \frac{\beta SI}{N} - \gamma I\\
$$
$$
\frac{dR}{dt} = \gamma I
$$

with $S$, $I$, and $R$ representing the number of sick, infected, and recovered individuals, $N$ is the total number of people in the system such that $S + I + R = N$ and $\gamma$ and $\beta$ are the rates (?) of recovery and infection, respectively.  

## Heteronormative SIR model for cooties
$$
\frac{dS_g}{dt} = \frac{-\beta_g S_g I_b}{N_b}
$$
$$
\frac{dI_g}{dt} = \frac{\beta_g S_g I_b}{N_b} - \gamma_g I_g\\
$$
$$
\frac{dR_g}{dt} = \gamma_g I_g
$$
$$
\frac{dS_b}{dt} = \frac{-\beta_b S_b I_g}{N_g}\\
$$
$$
\frac{dI_b}{dt} = \frac{\beta_b S_b I_g}{N_g} - \gamma_b I_b\\
$$
$$
\frac{dR_b}{dt} = \gamma_b I_b
$$

with $S_g$, $I_g$, and $R_g$ representing the number of sick, infected, and recovered girls and $S_b$, $I_b$, and $R_b$ representing the number of sick, infected, and recovered boys.  $N_g = S_g + I_g + R_g$ is the total number of girls, $N_b = S_b + I_b + R_b$ is the total number of boys, and $N = N_b + N_g$ is the total number of boys and girls in the population.  $\beta_g$ and $\beta_b$ are the infection rate(is it a rate?) for girls and boys, respectively.  Finally, $\gamma_g$ and $\gamma_b$ are the racovery rate(is it a rate?) for girls and boys, respectively. 


In the special case where $\beta_g = \beta_b$ and $\gamma_g = \gamma_b$, What happens.......

Proof: Is this even true?  Would be cool to have a proof!

$$
\frac{dS}{dt} = \frac{dS_g}{dt} + \frac{dS_b}{dt}
$$

$$
= \frac{-\beta S_g I_b}{N_b} + \frac{-\beta S_b I_g}{N_g} = -\beta\left(\frac{S_g I_b}{N_b} + \frac{S_b I_g}{N_g}\right)
$$


# Results {#sec:results}

# Conclusion {#sec:conclusion}

# Acknowledgements {-}

# Supplementary Material {-}

All code for reproducing the analyses in this paper is publicly available at https://github.com/gjm112/HockeyEnforcers

# References
