---
title: |
  |  \Large SIR cooties
author: | 
  | \large Gregory J. Matthews \vspace{-1.1mm}
  | \normalsize Department of Mathematics and Statistics \vspace{-1mm}
  | \normalsize Center for Data Science and Consulting \vspace{-1mm}
  | \normalsize Loyola University Chicago \vspace{-1mm}
  | \normalsize Chicago, IL 60660 \vspace{-1mm}
  | \normalsize [`gmatthews1@luc.edu`](mailto:gmatthews1@luc.edu) \vspace{-1mm}
abstract: |
  | Cooties. \vspace{2mm}
  | *Keywords*: Cooties
bibliography: references.bib
fontsize: 12pt
link-citations: true
linkcolor: cyan
urlcolor: cyan
output:
  pdf_document:
    df_print: kable
    number_sections: true
    keep_tex: true
header-includes:
 \usepackage{setspace}
 \setstretch{1.15}
 \usepackage{float}
 \floatplacement{figure}{t}
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE
)

gjm <- function(x, color = "red") {
  sprintf("\\textcolor{%s}{%s}", color, x)}
```

```{r pkgs}
library(tidyverse)
theme_set(theme_minimal())
library(Lahman)
```

\newpage

# Introduction {#sec:intro}
SIR models (@weiss2013sir) are commonly used as a simple model for epidemics.  

Covid: @ellison2020implications, @cooper2020sir, @kudryashov2021analytical

Measles: @bjornstad2002dynamics, @arsal2020short, @allen1991discrete


SIR Lit review

## Zombies
Many 

@reehl2013zombie, @simeonov2022humans, @witkowski2013bayesian, @libal2023transition, @alemi2015you

# Methods {#sec:methods}
These models can be expressed as differential equations with the solution describing the trajectories of the susceptible-infected-recovered groups.  The traditional SIR model is described by the following differential equations: 

$$
\frac{dS}{dt} = \frac{-\beta SI}{N}\\
$$
$$
\frac{dI}{dt} = \frac{\beta SI}{N} - \gamma I\\
$$
$$
\frac{dR}{dt} = \gamma I
$$

with $S$, $I$, and $R$ representing the number of sick, infected, and recovered individuals, $N$ is the total number of people in the system such that $S + I + R = N$ and $\gamma$ and $\beta$ are the rates (?) of recovery and infection, respectively.  

## Heteronormative SIR model for cooties
$$
\frac{dS_g}{dt} = \frac{-\beta_g S_g I_b}{N_b}
$$
$$
\frac{dI_g}{dt} = \frac{\beta_g S_g I_b}{N_b} - \gamma_g I_g\\
$$

$$
\frac{dR_g}{dt} = \gamma_g I_g
$$
$$
\frac{dS_b}{dt} = \frac{-\beta_b S_b I_g}{N_g}\\
$$
$$
\frac{dI_b}{dt} = \frac{\beta_b S_b I_g}{N_g} - \gamma_b I_b\\
$$
$$
\frac{dR_b}{dt} = \gamma_b I_b
$$

with $S_g$, $I_g$, and $R_g$ representing the number of sick, infected, and recovered girls and $S_b$, $I_b$, and $R_b$ representing the number of sick, infected, and recovered boys.  $N_g = S_g + I_g + R_g$ is the total number of girls, $N_b = S_b + I_b + R_b$ is the total number of boys, and $N = N_b + N_g$ is the total number of boys and girls in the population.  $\beta_g$ and $\beta_b$ are the infection rate(is it a rate?) for girls and boys, respectively.  Finally, $\gamma_g$ and $\gamma_b$ are the racovery rate(is it a rate?) for girls and boys, respectively. 


In the special case where $\beta_g = \beta_b$ and $\gamma_g = \gamma_b$, What happens.......

In some scribbling, I have that the models are equivalent if 
$$
(S_gN_g -S_bN_b)(I_bN_g - I_gN_b) = 0
$$
So if either of these are true: 
This first one is the infected ratio.  
$$
\frac{I_b}{N_b} = \frac{I_g}{N_g} 
$$
or (I don't think thi second one makes sense) 
$$
\frac{S_b}{N_g} = \frac{S_g}{N_b} 
$$

Proof: Is this even true?  Would be cool to have a proof!

$$
\frac{dS}{dt} = \frac{dS_g}{dt} + \frac{dS_b}{dt}
$$

$$
= \frac{-\beta S_g I_b}{N_b} + \frac{-\beta S_b I_g}{N_g} = -\beta\left(\frac{S_g I_b}{N_b} + \frac{S_b I_g}{N_g}\right)
$$

```{r}
Sg <- 100
Sb <- 50
Ib <- 3
Ig <- 3
Nb <- Sb + Ib
Ng <- Sg + Ig

(Sg + Sb)*(Ig+Ib)/ (Nb+Ng)

Sg*Ib/Nb + Sb*Ig/Ng
```


# Results {#sec:results}
Balanced
```{r}
Sb <- 150
Sg <- 150
Ib <- 1
Ig <- 1
Rb <- 0
Rg <- 0

library(deSolve)
closed.sir.model <- function (t, x, params) {
  ## first extract the state variables
  S <- x[1]
  I <- x[2]
  R <- x[3]
  ## now extract the parameters
  beta <- params["beta"]
  gamma <- params["gamma"]
  N <- S+I+R
  ## now code the model equations
  dSdt <- -beta*S*I/N
  dIdt <- beta*S*I/N-gamma*I
  dRdt <- gamma*I
  ## combine results into a single vector
  dxdt <- c(dSdt,dIdt,dRdt)
  ## return result as a list!
  list(dxdt)
}


parms <- c(beta=0.25,gamma=1/60)
times <- seq(from=0,to=180,by=1)
xstart <- c(S=Sg+Sb,I=Ib+Ig,R=Rb+Rg)


library(tidyverse)
ode(
  func=closed.sir.model,
  y=xstart,
  times=times,
  parms=parms
) %>%
  as.data.frame() -> out1



#out1 %>%
  # gather(variable,value,-time) %>%
  # ggplot(aes(x=time,y=value,color=variable))+
  # geom_line(size=2)+
  # theme_classic()+
  # labs(x='time (days)',y='number of individuals')
```

This model has infections peaking on day `r which.max(out1$I)`.  
Same starting conditions but balanced between the two groups.  
```{r}
#Now for cooties with recoveries
closed.sir.model <- function (t, x, params) {
  ## first extract the state variables
  Sg <- x[1]
  Sb <- x[2]
  Ig <- x[3]
  Ib <- x[4]
  Rg <- x[3]
  Rb <- x[4]
  ## now extract the parameters
  beta <- params["beta"]
  gamma <- params["gamma"]
  Nb <- Sb+Ib + Rb
  Ng <- Sb+Ig + Rg
  ## now code the model equations
  dSgdt <- -beta*Sg*Ib/Nb
  dSbdt <- -beta*Sb*Ig/Ng
  dIgdt <- beta*Sg*Ib/Nb - gamma*Ig
  dIbdt <- beta*Sb*Ig/Ng - gamma*Ib
  dRgdt <- gamma*Ig
  dRbdt <- gamma*Ib
  ## combine results into a single vector
  dxdt <- c(dSgdt,dSbdt,dIgdt,dIbdt,dRgdt,dRbdt)
  ## return result as a list!
  list(dxdt)
}


parms <- c(beta=0.25,gamma=1/60)
times <- seq(from=0,to=180,by=1)
xstart <- c(Sg=Sg, Sb=Sb,Ig=Ig, Ib = Ib, Rg = Rg, Rb=Rb)

library(tidyverse)
ode(
  func=closed.sir.model,
  y=xstart,
  times=times,
  parms=parms
) %>%
  as.data.frame() -> out


out <- out %>% mutate(S = Sg + Sb, I = Ib+Ig, R = Rb+Rg)



# out %>%
#   select(S, I, R, time) %>% 
#   gather(variable,value,-time) %>%
#   ggplot(aes(x=time,y=value,color=variable))+
#   geom_line(size=2)+
#   theme_classic()+
#   labs(x='time (days)',y='number of individuals') 

dat1 <- out1 %>% gather(variable,value,-time)
#Cooties model
dat <- out %>%select(S, I, R, time) %>% gather(variable,value,-time)

  ggplot()+
  geom_line(aes(x=time,y=value,color=variable), data = dat1,size=1, lty = 2)+  geom_line(aes(x=time,y=value,color=variable), data = dat ,size=1) + 
  theme_classic()+
  labs(x='time (days)',y='number of individuals')

```


```{r}
Sb <- 200
Sg <- 50
Ib <- 1
Ig <- 0
Rb <- 0
Rg <- 0

library(deSolve)
closed.sir.model <- function (t, x, params) {
  ## first extract the state variables
  S <- x[1]
  I <- x[2]
  R <- x[3]
  ## now extract the parameters
  beta <- params["beta"]
  gamma <- params["gamma"]
  N <- S+I+R
  ## now code the model equations
  dSdt <- -beta*S*I/N
  dIdt <- beta*S*I/N-gamma*I
  dRdt <- gamma*I
  ## combine results into a single vector
  dxdt <- c(dSdt,dIdt,dRdt)
  ## return result as a list!
  list(dxdt)
}


parms <- c(beta=0.25,gamma=1/60)
times <- seq(from=0,to=180,by=1)
xstart <- c(S=Sg+Sb,I=Ib+Ig,R=Rb+Rg)


library(tidyverse)
ode(
  func=closed.sir.model,
  y=xstart,
  times=times,
  parms=parms
) %>%
  as.data.frame() -> out1



#out1 %>%
  # gather(variable,value,-time) %>%
  # ggplot(aes(x=time,y=value,color=variable))+
  # geom_line(size=2)+
  # theme_classic()+
  # labs(x='time (days)',y='number of individuals')
```

This model has infections peaking on day `r which.max(out1$I)`.  
Same starting conditions but balanced between the two groups.  
```{r}
#Now for cooties with recoveries
closed.sir.model <- function (t, x, params) {
  ## first extract the state variables
  Sg <- x[1]
  Sb <- x[2]
  Ig <- x[3]
  Ib <- x[4]
  Rg <- x[3]
  Rb <- x[4]
  ## now extract the parameters
  beta <- params["beta"]
  gamma <- params["gamma"]
  Nb <- Sb+Ib + Rb
  Ng <- Sb+Ig + Rg
  ## now code the model equations
  dSgdt <- -beta*Sg*Ib/Nb
  dSbdt <- -beta*Sb*Ig/Ng
  dIgdt <- beta*Sg*Ib/Nb - gamma*Ig
  dIbdt <- beta*Sb*Ig/Ng - gamma*Ib
  dRgdt <- gamma*Ig
  dRbdt <- gamma*Ib
  ## combine results into a single vector
  dxdt <- c(dSgdt,dSbdt,dIgdt,dIbdt,dRgdt,dRbdt)
  ## return result as a list!
  list(dxdt)
}


parms <- c(beta=0.25,gamma=1/60)
times <- seq(from=0,to=180,by=1)
xstart <- c(Sg=Sg, Sb=Sb,Ig=Ig, Ib = Ib, Rg = Rg, Rb=Rb)

library(tidyverse)
ode(
  func=closed.sir.model,
  y=xstart,
  times=times,
  parms=parms
) %>%
  as.data.frame() -> out


out <- out %>% mutate(S = Sg + Sb, I = Ib+Ig, R = Rb+Rg)



# out %>%
#   select(S, I, R, time) %>% 
#   gather(variable,value,-time) %>%
#   ggplot(aes(x=time,y=value,color=variable))+
#   geom_line(size=2)+
#   theme_classic()+
#   labs(x='time (days)',y='number of individuals') 

dat1 <- out1 %>% gather(variable,value,-time)
#Cooties model
dat <- out %>%select(S, I, R, time) %>% gather(variable,value,-time)

  ggplot()+
  geom_line(aes(x=time,y=value,color=variable), data = dat1,size=1, lty = 2)+  geom_line(aes(x=time,y=value,color=variable), data = dat ,size=1) + 
  theme_classic()+
  labs(x='time (days)',y='number of individuals')

```




Because at any given time point only part of the population can infect the other part of the population, the overall infected rate grows slower than the the SIR model where everyone is capable of infecting everyone else.  

# Conclusion {#sec:conclusion}

# Acknowledgements {-}

# Supplementary Material {-}

All code for reproducing the analyses in this paper is publicly available at https://github.com/gjm112/HockeyEnforcers

# References
